Predicting political party affiliation from text

### 2016-06-22 manuscript/poltext.tex
polished. slowly converging, seemingly.

A major problem with automated text analyses is generalisation to text domains other than that on which the system has been trained~\cite{Slapin.2014}. 
Traditionally, experts conduct text analysis, who consider many contextual factors during the annotation of political text. There has, however, always been discussion about the reliability of such annotations. Automated analysis has the potential to solve this problem, but is more fragile with regard to questions of validity. 
%who judge and label political texts have read enough content of different domains to detect political bias in a variety of contexts and styles. 
% Den vorangegangenen Satz würde ich streichen, weil ich mit der Aussage nicht übereinstimme. Das ist wahrscheinlich wieder der Unterschied der Disziplinen, ihr müsst rechtfertigen, warum Computer es können, wenn Menschen es können, wir müssen in der Regel rechtgertigen warum Menschen es können, wenn Computer es besser machen. 
% Ich glaube, dass Computer es besser machen, wuerde niemand behaupten. Computer machen das genau so gut oder schlecht, wie die qualitaet der daten (die ja von menschen gelabeled oder selektiert werden) ihnen erlaubt. Der Algorithmus spiel auch ne rolle, aber das hauptproblem ist Daten und die koennen nur von Menschen kommen. Computer sind hilfreich als assistive technology, weil sie das ganze skalierbar machen. 
% Ich versteh deinen Kommentar, aber wir mussten uns tatsächlich am Anfang des Projekts viel mit so Fragen rumschlagen, warum wir denn noch mit Menschen kodieren, wenn Computer das doch besser können. Eigentlich geht es da aber, glaube ich, um das Spannugnsgeflecht zwischen Reliabilität und Validität. Habe versucht das jetzt nochmal so umzuformulieren. 
Machine learning algorithms are naturally prone to poor generalisation performance if the training data is biased towards one text domain. Unfortunately, good unbiased training data is difficult to obtain. One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data, as it consists of a large source of text that can be clearly associated with a party. We examine to what extent models trained on this data can be generalised to other text domains, such as party manifestos and texts from social media. 
We discuss the effects of text length and domain shift of text data, and investigate some potential reasons for the differences in classification performance.

@@

\subsection{Influence of text length on accuracy}
A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to classify, see also~\autoref{sec:data}. In order to investigate the effect of text length, we aggregated the data into longer texts, and group Manifesto data into political topics. \autoref{tab:results_topic} shows the topic level prediction results. We obtain F1 scores of above 0.8 for all parties except for the SPD . As the facebook posts lacked topic labels, we conducted the aggregation of these texts by sampling 50 texts of 1000 words from each of the concatenated facebook posts of a party. \ssc{this is a bit unclear, maybe another sentence with details would be good} The results are shown in \autoref{tab:results_fb}. Prediction accuracies comparable to the in-domain case can also be achieved for these texts. This increase is in line with previous findings on the influence of text length on political bias prediction accuracy~\cite{Hirst2014}.

@@

\subsection{Misclassifications and policy change}
An explanation for misclassifications could be that parties change their policy positions over time. This requires us to inspect the specific data points that were misclassified. The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that the SPD manifesto texts are often predicted as belonging to the CDU/CSU on the topic level. A similar effect can be observed in the 18th Bundestag, when we predict for both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU. One explanation for this misclassification could be policy changes. \ssc{That sounds too much like a speculation, can we maybe give one example, e.g. that the conservatives strengthened environmental issues?} \\

@@

\subsection{Discriminative features}
\label{sec:discrim_effect}
An important question is whether the difference between the features of each party stems from different policies or from other aspects of the text. In order to get more insight into what aspects are discriminative for government membership and for parties, we first analysed the effect of text sentiment and the textual features most correlated with each party. \ssc{this sentence is too long, but I don't have a good idea how to split it :D}

% ich glaub der uebergang von dem sentiment in der letzten subsection passt ganz gut zu den sentiment analysen im naechsten
% auch, weils beim sentiment noch um die binary classification geht, die gerade kam
\paragraph{Sentiment correlates with political power}
The drop in prediction accuracy in the government prediction task was more pronounced for manifesto texts than for facebook posts. What do facebook posts and plenary debates have in common? Both the speakers in the parliament as well as the authors of the facebook posts are aware of which party is in the government -- while the authors of the manifesto texts lack this knowledge, as these texts were written before the elections. A language feature that could capture this is sentiment. 
In order to assess this effect, we quantified the correlation between government membership as well as the number of seats in the parliament with text sentiment. Our results in \autoref{tab:sentiments} show that government membership correlates nearly perfect with positive sentiment, giving a correlation coefficient of 0.98, and the number of seats correlates with 0.89. Previous studies also find that text features discriminative in a two party system are not necessarily related to policies but more to language of defense and attack~\cite{Hirst2014}.

@@

In line with other findings \cite{Yu2008, Hirst2014}, we encounter a strong effect of text length and text domain on the generalization performance of the classifier. The first effect, that longer texts are easier to classify, intuitively makes sense. Humans are similarly worse at judging the political bias of shorter texts out of context. \cite{Benoit.Forthcoming}, for example, conducted an experiment where experts were differentiating whether specific sentences were dealing with economic or social policy. They observed only about 35\% agreement between all expert coders. However, short texts are a realistic \ssc{I don't understand this word here} challenge for automated political bias prediction systems: Political texts from social media data and other web media are indeed very short on many occasions. But political education can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions, even though they are difficult to analyse with human annotators due to the large volume of data. \\
The second effect, i.e. the drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments. In the case of party affiliation prediction, the out-of-domain classification is on a par or even better than the prediction accuracy on in-domain data. However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification. \ssc{sentence too long} We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.

>>>>

A major problem with automated text analyses is generalisation to text domains other than that on which the system has been trained~\cite{Slapin.2014}. 
%Traditionally, experts conduct text analysis, who consider many contextual factors during the annotation of political text. There has, however, always been discussion about the reliability of such annotations. Automated analysis has the potential to solve this problem, but is more fragile with regard to questions of validity. 
% Ich versteh deinen Kommentar, aber wir mussten uns tatsächlich am Anfang des Projekts viel mit so Fragen rumschlagen, warum wir denn noch mit Menschen kodieren, wenn Computer das doch besser können. Eigentlich geht es da aber, glaube ich, um das Spannugnsgeflecht zwischen Reliabilität und Validität. Habe versucht das jetzt nochmal so umzuformulieren. 
% hm. also validitaet is ja schon im abschnitt davor angesprochen - ich war ehrlich gesagt auch erstaunt, wie das labeling gemacht wurde. um validitaet abschaetzen zu koennen braucht man ja mindestens 3 labels pro datenpunkt. den paragraph mit generalisierung einzuleiten und dann die validitaet nochmal nur so anzureissen ohne darauf einzugehn wieso fand ich nich so fluessig. der eigentliche punkt hier war ja generalisierung, und da haben menschen den vorteil, dass manche menschen heterogenere samples gelesen haben im lauf ihres lebens als die algorithmen von uns kriegen. ich hab versucht das wieder klarer rauszuarbeiten.
While political experts have read texts of different domains and are able to detect political bias in a variety of contexts and styles, machine learning algorithms are prone to poor generalisation performance across text domains if the training data is biased towards one domain only. Unfortunately, good unbiased training data is difficult to obtain. One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data, as it consists of a large source of text that can be clearly associated with a party. We examine to what extent models trained on this data can generalise their predictions to other text domains, such as party manifestos and texts from social media. 
We discuss the effects of text length and domain shift of text data, and investigate some potential reasons for the differences in classification performance.

@@

\subsection{Influence of text length on accuracy}
A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to classify, see also~\autoref{sec:data}. In order to investigate the effect of text length, we aggregated the data into longer texts, and group Manifesto data into political topics. \autoref{tab:results_topic} shows the topic level prediction results. We obtain F1 scores of above 0.8 for all parties except for the SPD . As the facebook posts lacked topic labels, we conducted the aggregation of these texts by first concatenating all facebook posts of a party into one long text; this text was then partitioned into segments of 1000 words each. For each party 50 random segments were selected for classification. The results are shown in \autoref{tab:results_fb}. Prediction accuracies comparable to the in-domain case can also be achieved for these texts. This increase is in line with previous findings on the influence of text length on political bias prediction accuracy~\cite{Hirst2014}.

@@

\subsection{Misclassifications and policy change}
An explanation for misclassifications could be that parties change their policy positions over time. This requires us to inspect the specific data points that were misclassified. The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that the SPD manifesto texts are often predicted as belonging to the CDU/CSU on the topic level. A similar effect can be observed in the 18th Bundestag when predicting both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU. While we cannot list examples due to space limitations, we emphasize that these misclassifications can be related to policy changes of parties. \\
% there were concrete examples (17th bundestag social reforms -> SPD mixed up with CDU/CSU, 18th bundestag fukushima -> CDU/CSU mixed up with green party), which can not be included due to space constraints. we might leave this out then?

@@

\subsection{Discriminative features}
\label{sec:discrim_effect}
An important question is whether the difference between the features of each party stems from different policies or from other aspects of the text. To address this point we analysed features that are discriminative for government membership and for parties, respectively.

% ich glaub der uebergang von dem sentiment in der letzten subsection passt ganz gut zu den sentiment analysen im naechsten
% auch, weils beim sentiment noch um die binary classification geht, die gerade kam
\paragraph{Sentiment correlates with political power}
The drop in prediction accuracy in the government prediction task was more pronounced for manifesto texts than for facebook posts. What do facebook posts and plenary debates have in common? In contrast to the authors of manifestos both the speakers in the parliament as well as the authors of facebook posts know which party is in the government. A language feature that could capture this is sentiment. 
Indeed our results in \autoref{tab:sentiments} show that positive sentiment strongly correlates with government membership and the number of seats in the parliament.  Previous studies also find that text features discriminative in a two party system are not necessarily related to policies but more to language of defense and attack~\cite{Hirst2014}.

@@

In line with previous findings \cite{Yu2008, Hirst2014}, we encounter a strong effect of text length and text domain on the generalization performance of the classifier. The first effect, that longer texts are easier to classify, intuitively makes sense. Also humans are challenged when judging the political bias of shorter texts out of context \cite{Benoit.Forthcoming}.
%, for example, conducted an experiment where experts were differentiating whether specific sentences were dealing with economic or social policy. They observed only about 35\% agreement between all expert coders. 
% zu wenig platz fuer den satz, aber glaube, der punkt kommt auch so rueber: kurzer text is schwierig auf fuer menschen. 
% re challenge: das soll hier heissen, dass das zwar schwierig aber realistisch ist, weil mans ja machen muss in vielen kontexten. 
% ich hab versucht, das klarer zu machen, indem ich das in dem satz oben auch erwaehne. 
However, short texts are a realistic challenge for automated political bias prediction systems: Political texts from social media data and other web sources are indeed often very short and hence difficult to analyse for both human annotators and algorithms. Both political education and science can benefit from automatic analyses of these very data streams, as these media have strong influence on public opinions yet cannot be analysed by humans alone, due to the volume of data.

The second effect, i.e. the drop in generalization performance on out-of-domain data, appears to be correlated to the first one: it can be alleviated in some cases by aggregating texts into longer segments. In the case of party affiliation prediction, the out-of-domain classification is on a par or even better than the prediction accuracy on in-domain data. However in the binary classification setting (government membership prediction), text aggregation does not help as much: Aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain accuracy drop. We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
